---
title: "1_Statistical Summary"
output: html_document
date: "2022-05-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data reading using package **openxlsx**

```{r readingData}
require(openxlsx)
 mydata  <- read.xlsx("Data TSOs.xlsx", startRow=3)
 colnames(mydata)[ncol(mydata)] <- "Efficiency"
 mydata <- mydata[,-1]
 View(mydata)
 
 head(mydata)
 summary(mydata)
 names(mydata) # provide names of the columns 
 dim(mydata)   # provide data dimensionality 
 # Accessing specific columns
 #mydata$OPEX  # 1) this or the other ones
 #mydata[,1]      # 2) option 2
 #mydata[4:9,"OPEX"]  # 3) option 3 (rows 4 to 9 with the column  "Company")
 sd(mydata$OPEX)  #standard deviation
```

## Chunk 1: Visualization of OPEX & efficiency of all companies over the years

```{r GraphsBasics, fig.width=8}
# Data preparation
library("tidyverse")
df <-  mydata %>% select(year, company,OPEX, Efficiency) %>% slice(1:4, 7:10, 13:16, 19:22, 25:28, 31:34, 37:40, 43:46, 49:52, 55:74)
yearly_counts <- df %>% count(year, company, OPEX, Efficiency)

# Visualization of OPEX of all companies over the years
ggplot(data = yearly_counts, aes(x = year, y = OPEX, color = company)) +
  geom_line(size = 1.0) +
  scale_y_continuous(labels = scales::comma) + ylab("OPEX [$]") + xlab("Year") +
    theme(axis.title.y = element_text(size = 15, angle = 90, face = "bold")) +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(legend.title = element_text(size = 15, face="bold")) +
  theme(legend.text = element_text(size = 12))

  ##Pankaj Rajoria - 25.11.2022
  # theme(axis.title.y = element_text(size = 9, angle = 90, face = "bold")) +
  # theme(axis.title.x = element_text(size = 9, face = "bold")) +
  # theme(legend.title = element_text(size = 9, face="bold")) +
  # theme(legend.text = element_text(size = 9)) +
  # theme(axis.text = element_text(size = 9, face="bold")) +
  # theme(legend.key.size = unit(2, 'cm'), #change legend key size
  #       legend.key.height = unit(1, 'cm'), #change legend key height
  #       legend.key.width = unit(1, 'cm'), #change legend key width
  #       legend.title = element_text(face = "bold", size=9), #change legend title font size
  #       legend.text = element_text(size=9))

# Visualization of Efficiency.with.Weight.restrictions of all companies over the years
ggplot(data = yearly_counts, aes(x = year, y = Efficiency, color = company)) +
  geom_line(size = 1.3) + scale_y_continuous(labels = scales::comma) +
  ylab("Efficiency") +
  xlab("Year") + theme(axis.title.y = element_text(size = 9, angle = 90, face = "bold")) +
  theme(axis.title.y = element_text(size = 15, angle = 90, face = "bold")) +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(legend.title = element_text(size = 15, face="bold")) +
  theme(legend.text = element_text(size = 12))

  ##Pankaj Rajoria - 25.11.2022
  # theme(axis.title.x = element_text(size = 9, face = "bold")) +
  # theme(legend.title = element_text(size = 9, face="bold")) +
  # theme(legend.text = element_text(size = 9)) +
  # theme(axis.title.x = element_text(size = 9, face = "bold")) +
  # theme(legend.title = element_text(size = 9, face="bold")) +
  # theme(legend.text = element_text(size = 9))+
  # theme(axis.text = element_text(size = 9, face="bold")) +
  # theme(legend.key.size = unit(2, 'cm'), #change legend key size
  #       legend.key.height = unit(1, 'cm'), #change legend key height
  #       legend.key.width = unit(1, 'cm'), #change legend key width
  #       legend.title = element_text(face = "bold", size=9), #change legend title font size
  #       legend.text = element_text(size=9))


# Visualization of OPEX of all companies over the years with explicit marked companies
#  ggplot(data = yearly_counts, aes(x = OPEX, y = out, color = company), fill = "transparent") + ####Need to replace the variable y out with something else.
#   geom_point(size = 1.7) + scale_y_continuous(labels = scales::comma) +
#   ylab("Output") +
#   xlab("OPEX [$]") +
#   theme(axis.title.y = element_text(size = 15, angle = 90, face = "bold")) +
#   theme(axis.title.x = element_text(size = 15, face = "bold")) +
#   theme(legend.title = element_text(size = 15, face="bold")) +
#   theme(legend.text = element_text(size = 12)) 
```

## Chunk 2: Statistical summary for every attribute

```{r GraphsBasics, fig.width=8}
# Data preparation
library("tidyverse")
library("ggplot2")
library(ggpubr)
library("egg")
install.packages("png")
# library("ggsave")
df_2 <- mydata %>%
  select(company:Efficiency) %>% slice(1:4, 7:10, 13:16, 19:22, 25:28, 31:34, 37:40, 43:46, 49:52, 55:74)
  # dplyr::select(company:OPEX) %>% slice(1:4, 7:10, 13:16, 19:22, 25:28, 31:34, 37:40, 43:46, 49:52, 55:74)
yearly_counts_2 <- df_2 %>% count(OPEX, company, year)


# Statistical summary for every attribute
n <- 26
#This works only for i= 3,5,and 9 - 26
for(i in 26:n){
out <- df_2[, c(i)]
summary(df_2[, c(i)]) # Summary of the statistical data
var(df_2[, c(i)]) ## variance
sd(df_2[, c(i)]) ## standard deviation
data_summary <- round(summary(df_2[, c(i)]), digits = 2)
data_summary$Var <- round(var(df_2[, c(i)]), digits = 2)
data_summary$Sd <- round(sd(df_2[, c(i)]), digits = 2)

# Add the title to the first element of abc
  abc <- c(paste0(colnames(df_2[i])), unlist(data_summary))
  p_final <- ggtexttable(abc)
# abc <- lapply(data_summary, unlist)
# main_data <- as.data.frame(abc, stringsAsFactors = F)
# str(abc)

# p_final <- as.table(main_data)
# ggtexttable(main_data)

# first.step <- lapply(data_summary, unlist) 
# second.step <- as.data.frame(first.step, stringsAsFactors = F)
# rqtq <- ggtexttable(df, rows = NULL)

# Auto-Saving of the plots
# dev.off()
filename_Histo <- paste("SUMMARY_",colnames(df_2[i]),".png", sep = "")
# filename_Histo <- paste(yearly_counts2, "SUMMARY",".png", sep = "")
ggsave(device = "png", filename = filename_Histo, dpi = 400)

# png(filename=filename_Histo, width=960, height=960)
# ggsave(device = "png", filename = filename_Histo, width = 480, height = 480, units = "px")
# width = 480, height = 480, units = "px"
}
```

## Chunk 3: Plot of histogram & boxplot for every attribute

```{r GraphsBasics, fig.width=8}
# Data preparation
library("tidyverse")
library("ggplot2")
library("egg")
df_2 <- mydata %>%
  dplyr:: select(company:Efficiency) %>% slice(1:4, 7:10, 13:16, 19:22, 25:28, 31:34, 37:40, 43:46, 49:52, 55:74)
yearly_counts_2 <- df_2 %>% count(OPEX, company, year)
View(df_2)

# df_2 <- mydata %>%
#   select(company:Efficiency.without.Weight.restrictions) %>% slice(1:4, 7:10, 13:16, 19:22, 25:28, 31:34, 37:40, 43:46, 49:52, 55:74)
# yearly_counts_2 <- df_2 %>% count(OPEX, company, year)
# df_2

# Histogram & Boxplot for every attribute
n <- 26
for(i in 3:n){
out <- df_2[, c(i)]
p_1 <- ggplot(data = yearly_counts_2, aes(x = out, color = company)) + geom_histogram(aes(y =..density..),
                 col="black",
                 fill="red",
                 alpha=.2) +
  geom_density(col="black", size = 1.5, linetype = "solid") +
  labs(x=colnames(df_2[i]), y="Count") +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(labels = scales::comma) +
  theme(axis.title.y = element_text(size = 24, angle = 90, face = "bold")) +
  theme(axis.title.x = element_text(size = 24, face = "bold")) +
  theme(axis.text = element_text(color="black", face = "bold", size=15))

p_2 <- ggplot(data = yearly_counts_2, aes(x = out, color = company)) +
  geom_boxplot( col="black",
                 fill="red",
                 alpha=.2) +
  theme_classic() + scale_y_continuous(labels = scales::comma) + scale_x_continuous(labels = scales::comma) + labs(x=colnames(df_2[i])) +
  theme(axis.title.x = element_text(size = 24, face = "bold")) +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  theme(axis.text = element_text(color="black", face = "bold", size=15))

p_final <- egg::ggarrange(p_1, p_2, heights = 2:1)
#p_final <- ggarrange(p_1, p_2,
         #ncol = 1, nrow = 3,
         #heights = c(1, 0.5))
p_final

# Auto-Saving of the plots
# filename_Histo <- paste(colnames(df_2[i]), ".png", sep = "")
#ggsave(device = "png", filename = filename_Histo, dpi = 400)
png(file= "filename_Histo1.png",
width=1920, height=1080)
print(p_final)
dev.off()
}
```

## Chunk 4: Scatter plots of the outputs and enviromental variables

```{r GraphsBasics, fig.width=8}
# Data preparation
library("tidyverse")
library("ggplot2")
df_3 <- mydata %>%
 dplyr:: select(company:Efficiency) %>% slice(1:4, 7:10, 13:16, 19:22, 25:28, 31:34, 37:40, 43:46, 49:52, 55:74)
yearly_counts_3 <- df_3 %>% count(OPEX, company, year, Efficiency)
df_3



# Scatter Plot for the outputs
#only works for n = 9
n <- 9
for(i in n:25){
out_2 <- df_3[, c(i)]
p <- ggplot(data = yearly_counts_3, aes(x = out_2, y = OPEX, color = company)) + geom_point(size = 2.1) + scale_y_continuous(labels = scales::comma) + ylab(colnames(df_3[i])) + xlab("OPEX [$]") + theme(axis.title.y = element_text(size = 15, angle = 90, face = "bold")) + theme(axis.title.x = element_text(size = 15, face = "bold")) + theme(legend.title = element_text(size = 15, face="bold")) + theme(legend.text = element_text(size = 12))

# Auto-Saving of the plots
filename_Plot <- paste(colnames(df_3[i]), " Plot.png", sep = "")
ggsave(device = "png", filename = filename_Plot, dpi = 400)
}

# Scatter Plot for the enviromental variables
n <- 13
for(i in n:25){
out_3 <- df_3[, c(i)]
p <- ggplot(data = yearly_counts_3, aes(x = out_3, y = Efficiency.with.Weight.restrictions, color = company)) + geom_point(size = 2.1) + scale_y_continuous(labels = scales::comma) + ylab(colnames(df_3[i])) + xlab("Efficiency.with.Weight.restrictions") + theme(axis.title.y = element_text(size = 15, angle = 90, face = "bold")) + theme(axis.title.x = element_text(size = 15, face = "bold")) + theme(legend.title = element_text(size = 15, face="bold")) + theme(legend.text = element_text(size = 12))

filename_Plot <- paste(colnames(df_3[i]), " Plott.png", sep = "")
ggsave(device = "png", filename = filename_Plot, dpi = 400)
}
```

## Chunk 5: Boxplots for Regression-Chapter

```{r GraphsBasics, fig.width=8}
# Data preperation
df_kira <- mydata %>% dplyr::select(OPEX, Efficiency.with.Weight.restrictions, Efficiency.without.Weight.restrictions)

for (i in 1:3) {
  out <- df_kira[, c(i)]

  # Plot of regression Boxplots
  p_kira <- ggplot(data = df_kira, aes(x = out, color = company)) +
  geom_boxplot( col="black", width=0.7, fill="red", alpha=.2) +
  coord_flip() +
  theme_classic() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(labels = scales::comma) +
  labs(x=colnames(df_kira[i])) +
  theme(axis.title.x = element_text(size = 20, face = "bold")) +
  ylab("") +
  theme(axis.title.y = element_text(size = 20, face = "bold")) +
  theme(axis.text = element_text(face="bold")) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

print(p_kira)
filename_Histo <- paste(colnames(df_kira[i]), "KIRA",".png", sep = "")
ggsave(device = "png", filename = filename_Histo, width = 5, height = 10,dpi = 300)
}
```

## Chunk 6: Illustration of the relation between the OPEX and outputs/environmental variables

```{r GraphsBasics, fig.width=9}
# index of columns, start with Output 1
index <- 5

# go through all columns from Output 1 to Efficiency without weight restrictions
for (index in 5:28){

filename <- colnames(mydata)[index]
filename <- gsub("<", "lt", filename)
filename <- gsub(":", "", filename)
filename <- paste("OPEX_vs_",filename,".png", sep="")

# start export png
png(filename=filename, width=960, height=480)

# space for the plot
par(mfrow=c(1,1))
# space for the legend
par(mar=c(6,6,6,12), xpd=TRUE)

# package
require(car)

# subsets without artificially generated data and holdings
# OPEX subset
OPEX_subset <- mydata$OPEX[TSOs]
# outputs and environmental variables subsets
variable_subset <- mydata[,index][TSOs]

# declaration of x and y
x <- OPEX_subset
y <- variable_subset

# create dataframe for grouping by company name and year
df <- data.frame(x = x, y = y, group_name = factor(group), group_20XX=factor(group_year))

# plot and color by group and shape by year
plot(df$x, df$y, main =  paste("OPEX vs.", colnames(mydata)[index]), xlab = "OPEX", ylab= colnames(mydata)[index], col = colors[df$group_name], pch = pchs[df$group_20XX], cex = 2, frame = TRUE)

# create legends
legend(x="topright",inset=(c(-0.22,0)), y= "right", legend=c("CEEE-GT", "CELG G&T", "CEMIG-GT",  "CHESF", "COPEL-GT","CTEEP", "ELETRONORTE", "ELETROSUL","FURNAS", "STATE GRID_HOL", "TAESA_HOL", "TBE_HOL", "ALUPAR_HOL", "CELEO_HOL" ), fill=colors)

legend(x="bottomright", inset=(c(-0.22,0)), y= "right", legend = c("2013", "2014", "2015", "2016"), pch = pchs)

# stop export png
dev.off()

# increase index
index <- index+1
}
```

## Chunk 7: Illustration of the relation between each output/environmental variable and the efficiency without weight restrictions

```{r GraphsBasics, fig.width=9}
# index of columns, start with Output 1
index <- 5

# go through all columns from Output 1 to Environmental Variable 13
for (index in 5:26){

filename <- colnames(mydata)[index]
filename <- gsub("<", "lt", filename)
filename <- gsub(":", "", filename)
filename <- paste("C:\\Users\\Kira\\OneDrive\\Controlling Seminar\\Seminar Controlling\\Abbildungen\\output_environmental variables_vs_efficiency_without_wr\\Efficiency_wo_vs_",filename,".png", sep="")

# start export png
png(filename=filename, width=960, height=480)

# space for the plot
par(mfrow=c(1,1))
# space for the legend
par(mar=c(6,6,6,12), xpd=TRUE)

# package
require(car)

# subsets without artificially generated data and holdings
# Efficiency subset
efficiency_subset <- mydata$Efficiency.without.Weight.restrictions[TSOs]
# outputs and environmental variables subsets
variable_subset <- mydata[,index][TSOs]

# declaration of x and y
x <- variable_subset
y <- efficiency_subset

# create dataframe for grouping by company name and year
df <- data.frame(x = x, y = y, group_name = factor(group), group_20XX=factor(group_year))

# plot and color by group and shape by year
plot(df$x, df$y, main =  paste(colnames(mydata)[index], "vs. Efficiency without weight restrictions"), xlab = colnames(mydata)[index], ylab= "Efficiency without weight restrictions" , col = colors[df$group_name], pch = pchs[df$group_20XX], cex = 2,frame = TRUE)

# create legends
legend(x="topright",inset=(c(-0.22,0)), y= "right", legend=c("CEEE-GT", "CELG G&T", "CEMIG-GT",  "CHESF", "COPEL-GT","CTEEP", "ELETRONORTE", "ELETROSUL","FURNAS", "STATE GRID_HOL", "TAESA_HOL", "TBE_HOL", "ALUPAR_HOL", "CELEO_HOL" ), fill=colors)
legend(x="bottomright", inset=(c(-0.22,0)), y= "right", legend = c("2013", "2014", "2015", "2016"), pch = pchs)

# stop export png
dev.off()

# increase index
index <- index+1
}
```

## Chhunk 8: Illustration of the relation between each output/environmental variable and the efficiency with weight restrictions

```{r GraphsBasics, fig.width=9}
# index of columns, start with Output 1
index <- 5

# go through all columns from Output 1 to Environmental Variable 13
for (index in 5:26){

filename <- colnames(mydata)[index]
filename <- gsub("<", "lt", filename)
filename <- gsub(":", "", filename)
filename <- paste("C:\\Users\\Kira\\OneDrive\\Controlling Seminar\\Seminar Controlling\\Abbildungen\\output_environmental variables_vs_efficiency_with_wr\\Efficiency_w_vs_",filename,".png", sep="")

# start export png
png(filename=filename, width=960, height=480)

# space for the plot
par(mfrow=c(1,1))
# space for the legend
par(mar=c(6,6,6,12), xpd=TRUE)

# package
require(car)

# subsets without artificially generated data and holdings
# Efficiency subset
efficiency_subset <- mydata$Efficiency.with.Weight.restrictions[TSOs]
# outputs and environmental variables subsets
variable_subset <- mydata[,index][TSOs]

# declaration of x and y
x <- variable_subset
y <- efficiency_subset

# create dataframe for grouping by company name and year
df <- data.frame(x = x, y = y, group_name = factor(group), group_20XX=factor(group_year))

# plot and color by group and shape by year
plot(df$x, df$y, main =  paste(colnames(mydata)[index], "vs. Efficiency with weight restrictions"), xlab = colnames(mydata)[index], ylab= "Efficiency with weight restrictions" , col = colors[df$group_name], pch = pchs[df$group_20XX], cex = 2,frame = TRUE)

# create legends
legend(x="topright",inset=(c(-0.22,0)), y= "right", legend=c("CEEE-GT", "CELG G&T", "CEMIG-GT",  "CHESF", "COPEL-GT","CTEEP", "ELETRONORTE", "ELETROSUL","FURNAS", "STATE GRID_HOL", "TAESA_HOL", "TBE_HOL", "ALUPAR_HOL", "CELEO_HOL" ), fill=colors)
legend(x="bottomright", inset=(c(-0.22,0)), y= "right", legend = c("2013", "2014", "2015", "2016"), pch = pchs)

# stop export png
dev.off()

# increase index
index <- index+1
}
```
