---
title: "4_Regression_OPEX_Outputs"
output: html_document
date: '2022-07-18'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data reading using package **openxlsx**
```{r readingData}
require(openxlsx)
 mydata  <- read.xlsx("Data - Seminar.xlsx", startRow=3)

```
## chunk 1: Linear Regression model: OPEX ~ Outputs (identification of Costdrivers)
```{r}
# Pearson and Spearman correlation coefficients show strong positive correlation
# partial correlation analysis seems to indicate no direct correlations 
# scatterplots show that a linear regression might be suitable 
```
## chunk 2: definition of subset 
```{r}
# subset without artificially generated data for balanced plm 
companies_raw <- c(-5,-6,-11,-12,-17,-18,-23,-24,-29,-30,-35,-36,-41,-42,-47,-48,-53, -54)
mydata_balanced <- mydata[companies_raw, -1]

```
## chunk 3: OLS model: OPEX ~ outputs

```{r}
require (robustbase)
require(car)

OLS <-lm(formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced)

summary(OLS)

## Variance Inflation Factor Analysis
# to check if the variables have correlation or not. For example, cons: 0.98. That mean 98% of "cons" can be explained by the other variables. So, we may exclude this variable if needed. When, "NTLaj" is 0.54, that means only 54% of this variable can be explained by the other variables.

print(vif(OLS))
print(1 - 1/vif(OLS))

# if VIF is higher than 1/(1-R_squared), where R_squared is the multiple R_squared, then the predictor is more related to other predictors than to the response 
print("1/(1-R_squared)")
print(1/(1-0.9654))

# print residualplot of OLC 
residualPlot(OLS)

```
## chunk 4: OLS model: OPEX ~ outputs with feature selection 
```{r}
require(robustbase)
require(car)
OLS <- lm (formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced)

# Stepwise forward and backward selection
OLS <- step(OLS,trace = 0, k = 2, direction = "both")

summary(OLS)

# print the calculated AIC of OLS 
print("AIC")
print(AIC(OLS, k=2))


# OLS after feature selection adjusted according to VIF

OLS_adj <- lm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV` + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced)

summary (OLS_adj)

print("AIC OLS_adj")
print(AIC(OLS_adj, k=2))

# Breusch-Pagan test for homoskedasticity in a Linear Regression Model 

require(lmtest)

BP <- bptest(OLS)

print(BP)


## Variance Inflation Factor Analysis
# to check if the variables have correlation or not. For example, cons: 0.98. That mean 98% of "cons" can be explained by the other variables. So, we may exclude this variable if needed. When, "NTLaj" is 0.54, that means only 54% of this variable can be explained by the other variables.

# OLS
print(vif(OLS))
print(1 - 1/vif(OLS))
print("1/(1-R2)")
print(1/(1-0.9627))


# OLS_adj
print("OLS_adj")
print(vif(OLS_adj))
print(1 - 1/vif(OLS_adj))
print("1/(1-R2)")
print(1/(1-0.9608))


# Breusch-Pagan test for homoskedasticity 

require(lmtest)

BP <- bptest(OLS_adj)

print(BP)


# print residualplot of OLS
residualPlot(OLS_adj)

# model prediction
yhat <- predict(OLS_adj)

# compute MSE 
require(ModelMetrics)

MSE <- mse(mydata_balanced$OPEX, yhat)
print("MSE")
print(MSE)

#compute RMSE
RMSE <- rmse(mydata_balanced$OPEX, yhat)
print("RMSE")
print(RMSE)

```
## chunk 5: cross-validation OLS_adj model after feature selection 
```{r}

require(caret)
set.seed(123)

# cross validation 

# Define training control
 train.control <- trainControl(method = "cv", number = 10)

# Internal error in the train function is caused by using the model formula. Addressing the variables by their index avoided that error. 
 
# Train the model
 model <- train(x = mydata_balanced[,c(7,8,11)], y = mydata_balanced[,3], method = "lm", trControl = train.control)
 
# Summarize the results
 summary(model)

```

## chunk 6: Pooled OLS model: OPEX ~ outputs after feature selection 

```{r}
require(plm)

POLS <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced, model = "pooling", index = NULL)

summary (POLS)

# the Pooled OLS model of panel data is equal to the classic OLS model

```
## chunk 7: check assumption of homoscedasticity for Pooled OLS model: OPEX ~ outputs after feature selection 
```{r}
# space for the plot
par(mfrow=c(1,1))

# residuals of OLS model
res <- as.vector(OLS$residuals)
print(res)
# fitted values from the OLS model 
fitval <- as.vector(fitted(OLS))
print(fitval)

# plot residuals against fitted values 
plot(fitval, res, xlab = "fitted value", ylab = "residuals", main = "Homoskedasticity Test")
abline(lm(res ~ fitval, data = data.frame(res, fitval)), col = "blue")


# Breusch-Pagan test for homoskedasticity 

require(lmtest)

BP <- bptest(OLS)

print(BP)

# Durbin-Watson-Test for autocorrelation of disturbances 
# The Durbin-Watson-Test will have one output between 0 – 4. The mean (= 2) would indicate that there is no autocorrelation identified, 0 – 2 means positive autocorrelation (the nearer to zero the higher the correlation), and 2 – 4 means negative autocorrelation (the nearer to four the higher the correlation).

dwtest(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced)

```
## chunk 8: Pooled OLS model: OPEX ~ outputs
```{r}
require(plm)

OLS <- plm(formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced,  model = "pooling", index=NULL, row.names=TRUE)

summary(OLS)

```
## chunk 9: check assumption of homoscedasticity for Pooled OLS model: OPEX ~ outputs 
```{r}
# space for the plot
par(mfrow=c(1,1))

# residuals of OLS model
res <- as.vector(OLS$residuals)
print(res)
# fitted values from the OLS model 
fitval <- as.vector(fitted(OLS))
print(fitval)

# plot residuals against fitted values 
plot(fitval, res, xlab = "fitted value", ylab = "residuals", main = "Homoskedasticity Test")
abline(lm(res ~ fitval, data = data.frame(res, fitval)), col = "blue")


# Breusch-Pagan test for homoskedasticity 

require(lmtest)

BP <- bptest(OLS)

print(BP)

```
## chunk 10: check assumption of non-autocorrelation for Pooled OLS model: Opex ~ outputs 
```{r}
require (lmtest)

# Durbin-Watson-Test for autocorrelation of disturbances 

dwtest(formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced)



```
## chunk 11: Fixed Effect Model (FE): OPEX ~ outputs after feature selection 
```{r}

require(plm)


# FE with individual effects 

FE_individual <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced, model = "within", index = NULL, effect = "individual")

summary(FE_individual)
print("fixed effects")
summary(fixef(FE_individual))

# FE with time effects 

FE_time <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced, model = "within", index = NULL, effect = "time")

summary(FE_time)
print("fixed effects")
summary(fixef(FE_time))


# FE with both effects 

FE_both <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced, model = "within", index = NULL, effect = "twoways")

summary(FE_both)
print("fixed effects")
summary(fixef(FE_both))

# model prediction
yhat_individual <- predict(FE_individual)
yhat_time <- predict(FE_time)
yhat_both <- predict(FE_both)

# compute MSE 
require(ModelMetrics)

print("MSE individual")
print(mse(mydata_balanced$OPEX, yhat_individual))
print("MSE time")
print(mse(mydata_balanced$OPEX, yhat_time))
print("MSE both")
print(mse(mydata_balanced$OPEX, yhat_both))

#compute RMSE

print("RMSE individual")
print(rmse(mydata_balanced$OPEX, yhat_individual))
print("RMSE time")
print(rmse(mydata_balanced$OPEX, yhat_time))
print("RMSE both")
print(rmse(mydata_balanced$OPEX, yhat_both))



```
## chunk 12: F-Test for fixed effects 
```{r}

# H0: no difference between the FE and the Pooled model 
# reject the hypothesis if p-value < 0.05 

# individual effects 
pFtest(FE_individual, OLS_adj)


# time effects 
pFtest(FE_time, OLS_adj)

# both effects
pFtest(FE_both, OLS_adj)


```

## chunk 13: Random Effect Model (RE): OPEX ~ outputs after feature selection
```{r}
require(plm)

# RE with individual effects

RE_individual <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced, model = "random", index = NULL, effect = "individual")

summary(RE_individual)

# RE with  time effects 

RE_time <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced, model = "random", index = NULL, effect = "time", random.method = "amemiya")

summary(RE_time)

# RE with  individual and time effects 

RE_both <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + 
    mydata_balanced$`Number.of.switch.modes.<.230.kV`  + 
    mydata_balanced$`Reactive.power.(Mvar)`, data = mydata_balanced, model = "random", index = NULL, effect = "twoways", random.method = "amemiya")

summary(RE_both)

# model prediction
yhat_individual <- predict(RE_individual)
yhat_time <- predict(RE_time)
yhat_both <- predict(RE_both)

# compute MSE 
require(ModelMetrics)

print("MSE individual")
print(mse(mydata_balanced$OPEX, yhat_individual))
print("MSE time")
print(mse(mydata_balanced$OPEX, yhat_time))
print("MSE both")
print(mse(mydata_balanced$OPEX, yhat_both))

#compute RMSE

print("RMSE individual")
print(rmse(mydata_balanced$OPEX, yhat_individual))
print("RMSE time")
print(rmse(mydata_balanced$OPEX, yhat_time))
print("RMSE both")
print(rmse(mydata_balanced$OPEX, yhat_both))


```

# chunk 14: Lagrange-Multiplier-Test  for random effects 
```{r}
# H0: effects cannot be modeled as random variables 
# Ho is rejected, if p-value < 0.05

# individual effects 
plmtest(RE_individual, effect = "individual")


# time effects 
plmtest(RE_time, effect = "time")

# both effects
plmtest(RE_both, effect = "twoways")

```
## chunk 15: Hausman Test to compare FE and RE after feature selection 
```{r}
# The null hypothesis of this test states that random effects is preferred over fixed effects
# Ho: there is no correlation between the subject-specific error term and the independent variables

phtest(FE_individual, RE_individual)
phtest(FE_time, RE_time)
phtest(FE_both, RE_both)


```
## chunk 16:  FE Model: OPEX ~ outputs
```{r}
# FE model individual effects 
FE_individual <- plm (formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced, model = "within", index = NULL, effect = "individual")

summary(FE_individual)

# FE model time effects 
FE_time <- plm (formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced, model = "within", index = NULL, effect = "time")

summary(FE_time)

# FE model both effects 
FE_both <- plm (formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced, model = "within", index = NULL, effect = "twoways")

summary(FE_both)

```
## chunk 17:  F-Test for fixed effects 
```{r}

# H0: no difference between the FE and the Pooled model 
# reject the hypothesis if p-value < 0.05 

# individual effects 
pFtest(FE_individual, OLS)


# time effects 
pFtest(FE_time, OLS)

# both effects
pFtest(FE_both, OLS)

```
## chunk 18: RE Model: OPEX ~ outputs
```{r}
# RE model individual effects 
RE_individual <- plm (formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced, model = "random", index = NULL, effect = "individual")

summary(RE_individual)

# RE model time effects 
RE_time <- plm (formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced, model = "random", index = NULL, effect = "time", random.method = "amemiya")

summary(RE_time)

# RE model both effects 
RE_both <- plm (formula= mydata_balanced$OPEX ~ mydata_balanced$`Network.extension.<.230.kV` + mydata_balanced$`Network.extension.≥.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.<.230.kV` + mydata_balanced$`Number.of.power.transformers.and.reactors.≥.230.kV` + mydata_balanced$`Number.of.switch.modes.<.230.kV` + mydata_balanced$`Number.of.switch.modes.≥.230.kV` + mydata_balanced$`Transformation.capacity.(MVA)` + mydata_balanced$`Reactive.power.(Mvar)` + mydata_balanced$`Quality.variable:.Power.not.delivered`, data = mydata_balanced, model = "random", index = NULL, effect = "twoways", random.method = "amemiya")

summary(RE_both)
```
# chunk 19: Lagrange-Multiplier-Test  for random effects 
```{r}
# H0: effects cannot be modeled as random variables 
# Ho is rejected, if p-value < 0.05

# individual effects 
plmtest(RE_individual, effect = "individual")


# time effects 
plmtest(RE_time, effect = "time")

# both effects
plmtest(RE_both, effect = "twoways")

```
## chunk 20: Hausman Test to compare FE and RE after feature selection 
```{r}
# The null hypothesis of this test states that random effects is preferred over fixed effects
# Ho: there is no correlation between the subject-specific error term and the independent variables

phtest(FE_individual, RE_individual)
phtest(FE_time, RE_time)
phtest(FE_both, RE_both)

```

