---
title: "5_Regression_OPEX_Environmentals"
output: html_document
date: '2022-07-18'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Data reading using package **openxlsx**
```{r readingData}
require(openxlsx)
 mydata  <- read.xlsx("Data - Seminar.xlsx", startRow=3)

```
## chunk 1:  Linear Regression model: OPEX ~ environmental variables (identification of costdrivers/location factors)
```{r}
# Pearson and Spearman correlation coefficients show no significant correlation 
# partial correlation analysis seems to indicate no direct correlations 
# scatterplots show that a linear regression might be suitable 
```
## chunk 2: definition of subset 
```{r}
# subset without artificially generated data for balanced plm 
companies_raw <- c(-5,-6,-11,-12,-17,-18,-23,-24,-29,-30,-35,-36,-41,-42,-47,-48,-53, -54)
mydata_balanced <- mydata[companies_raw, -1]

```
## chunk 3: OLS model: OPEX ~ environmental variables 

```{r}
require (robustbase)
require(car)

OLS <-lm(formula= mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets + mydata_balanced$`(Age.of.assets).square` + mydata_balanced$Average.declivity + mydata_balanced$Intersection.with.indigenous.areas + mydata_balanced$Height.of.vegetation + mydata_balanced$Percentage.of.high.vegetation + mydata_balanced$Average.precipitation + mydata_balanced$Network.density + mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + mydata_balanced$Number.of.access.to.roads, data = mydata_balanced)

summary(OLS)

## Variance Inflation Factor Analysis
# to check if the variables have correlation or not. For example, cons: 0.98. That mean 98% of "cons" can be explained by the other variables. So, we may exclude this variable if needed. When, "NTLaj" is 0.54, that means only 54% of this variable can be explained by the other variables.

print(vif(OLS))
print(1 - 1/vif(OLS))

# if VIF is higher than 1/(1-R_squared), where R_squared is the multiple R_squared, then the predictor is more related to other predictors than to the response 
print("1/(1-R_squared)")
print(1/(1-0.9826))

# print residualplot of OLS 
residualPlot(OLS)

```
## chunk 4: OLS model: OPEX ~ environmental variables  with feature selection 
```{r}
require(robustbase)
require(car)
OLS <- lm (formula= mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets + mydata_balanced$`(Age.of.assets).square` + mydata_balanced$Average.declivity + mydata_balanced$Intersection.with.indigenous.areas + mydata_balanced$Height.of.vegetation + mydata_balanced$Percentage.of.high.vegetation + mydata_balanced$Average.precipitation + mydata_balanced$Network.density + mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + mydata_balanced$Number.of.access.to.roads, data = mydata_balanced)

# Stepwise forward and backward selection
OLS <- step(OLS,trace = 0, k = 2, direction = "both")

summary(OLS)

# print the calculated AIC of OLS 
print("AIC")
print(AIC(OLS, k=2))

# OLS after feature selection adjusted according to VIF

OLS_adj <- lm(formula =  mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets + 
    mydata_balanced$Average.declivity + 
    mydata_balanced$Height.of.vegetation  + 
    mydata_balanced$Average.precipitation + mydata_balanced$Network.density + 
    mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + 
    mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + 
    mydata_balanced$Number.of.access.to.roads, data = mydata_balanced)

summary (OLS_adj)

print("AIC OLS_adj")
print(AIC(OLS_adj, k=2))


# Breusch-Pagan test for homoskedasticity

require(lmtest)

BP <- bptest(OLS_adj)

print(BP)

# print residualplot of OLS
residualPlot(OLS_adj)

## Variance Inflation Factor Analysis
# to check if the variables have correlation or not. For example, cons: 0.98. That mean 98% of "cons" can be explained by the other variables. So, we may exclude this variable if needed. When, "NTLaj" is 0.54, that means only 54% of this variable can be explained by the other variables.

# OLS
print(vif(OLS))
print(1 - 1/vif(OLS))
print("1/(1-R_suqared")
print(1/(1-0.9825))

print("OLS_adj")
print(vif(OLS_adj))
print(1 - 1/vif(OLS_adj))
print("1/(1-R_suqared")
print(1/(1-0.9733))

# model prediction
yhat <- predict(OLS_adj)

# compute MSE 
require(ModelMetrics)

MSE <- mse(mydata_balanced$OPEX, yhat)
print("MSE")
print(MSE)

#compute RMSE
RMSE <- rmse(mydata_balanced$OPEX, yhat)
print("RMSE")
print(RMSE)



```
## chunk 5: cross-validation lm model after feature selection 
```{r}

require(caret)
set.seed(123)

# cross validation

# Define training control
 train.control <- trainControl(method = "cv", number = 10)

# Train the model
 model <- train(x = mydata_balanced[,c(13, 15, 17, 19, 20, 21, 22, 23, 24, 25)], y = mydata_balanced[,3], method = "lm", trControl = train.control)
 
# Summarize the results
 summary(model)

```
## chunk 6: Pooled OLS model: OPEX ~ environmental variables after feature selection 

```{r}
require(plm)

POLS <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets + mydata_balanced$Average.declivity + 
    mydata_balanced$Height.of.vegetation + 
    mydata_balanced$Average.precipitation + mydata_balanced$Network.density + 
    mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + 
    mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + 
    mydata_balanced$Number.of.access.to.roads, data = mydata_balanced, model = "pooling", index = NULL)

summary (POLS)

```
## chunk 7: check assumption of homoscedasticity and auto-correlation for OLS model: OPEX ~ environmental variables after feature selection 
```{r}
# space for the plot
par(mfrow=c(1,1))

# residuals of OLS model
res <- as.vector(OLS$residuals)

# fitted values from the OLS model 
fitval <- as.vector(fitted(OLS))


# plot residuals against fitted values 
plot(fitval, res, xlab = "fitted value", ylab = "residuals", main = "Homoskedasticity Test")
abline(lm(res ~ fitval, data = data.frame(res, fitval)), col = "blue")

# Breusch-Pagan test for homoskedasticity in a Linear Regression Model 

require(lmtest)

BP <- bptest(OLS)

print(BP)

# Durbin-Watson-Test for autocorrelation of disturbances 

dwtest(formula = mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets  + mydata_balanced$Average.declivity + 
    mydata_balanced$Height.of.vegetation +  
    mydata_balanced$Average.precipitation + mydata_balanced$Network.density + 
    mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + 
    mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + 
    mydata_balanced$Number.of.access.to.roads, data = mydata_balanced)

```
## chunk 8: Fixed Effect Model (FE): OPEX ~ outputs after feature selection 
```{r}

require(plm)


# FE with individual effects 

FE_individual <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets + 
    mydata_balanced$Average.declivity + mydata_balanced$Height.of.vegetation + 
    mydata_balanced$Average.precipitation + mydata_balanced$Network.density + 
    mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + 
    mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + 
    mydata_balanced$Number.of.access.to.roads, data = mydata_balanced, 
    model = "within")


summary(FE_individual)
print("fixed effects")
summary(fixef(FE_individual))

# FE with time effects 

FE_time <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets + mydata_balanced$Average.declivity + 
    mydata_balanced$Height.of.vegetation + 
    mydata_balanced$Average.precipitation + mydata_balanced$Network.density + 
    mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + 
    mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + 
    mydata_balanced$Number.of.access.to.roads, data = mydata_balanced, model = "within", index = NULL, effect = "time")

summary(FE_time)
print("fixed effects")
summary(fixef(FE_time))


# FE with both effects 

FE_both <- plm(formula = mydata_balanced$OPEX ~ mydata_balanced$Age.of.assets + mydata_balanced$Average.declivity + 
    mydata_balanced$Height.of.vegetation + 
    mydata_balanced$Average.precipitation + mydata_balanced$Network.density + 
    mydata_balanced$Fire.incidence + mydata_balanced$Number.of.intersections.with.roads + 
    mydata_balanced$Coincidence.of.network.lines + mydata_balanced$Density.of.lightening + 
    mydata_balanced$Number.of.access.to.roads, data = mydata_balanced, model = "within", index = NULL, effect = "twoways")

summary(FE_both)
print("fixed effects")
summary(fixef(FE_both))

# model prediction
yhat_individual <- predict(FE_individual)
yhat_time <- predict(FE_time)
yhat_both <- predict(FE_both)

# compute MSE 
require(ModelMetrics)

print("MSE individual")
print(mse(mydata_balanced$OPEX, yhat_individual))
print("MSE time")
print(mse(mydata_balanced$OPEX, yhat_time))
print("MSE both")
print(mse(mydata_balanced$OPEX, yhat_both))

#compute RMSE

print("RMSE individual")
print(rmse(mydata_balanced$OPEX, yhat_individual))
print("RMSE time")
print(rmse(mydata_balanced$OPEX, yhat_time))
print("RMSE both")
print(rmse(mydata_balanced$OPEX, yhat_both))


```
## chunk 9:  F-Test for fixed effects 
```{r}

# H0: no difference between the FE and the Pooled model 
# reject the hypothesis if p-value < 0.05 

# individual effects 
pFtest(FE_individual, OLS_adj)


# time effects 
pFtest(FE_time, OLS_adj)

# both effects
pFtest(FE_both, OLS_adj)


```

