---
title: "9_Poly_Regression_Efficiency_with_wr_Environmentals"
output: html_document
date: '2022-07-18'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data reading using package **openxlsx**
```{r readingData}
require(openxlsx)
 mydata  <- read.xlsx("Data - Seminar.xlsx", startRow=3)

```
## chunk 1: Polynomial Regression model: efficiency with weight restrictions ~ environmental variables (identification of costdrivers/location factors)
```{r}
# Pearson and Spearman correlation coefficients show weak significant correlation 
# partial correlation analysis seems to indicate weak direct correlations 
# scatterplots show that a linear regression might not be suitable 
```
## chunk 2: definition of subset 
```{r}
# subset without artificially generated data for balanced plm 
companies_raw <- c(-5,-6,-11,-12,-17,-18,-23,-24,-29,-30,-35,-36,-41,-42,-47,-48,-53, -54)
mydata_balanced <- mydata[companies_raw, -1]

```
## chunk 3: polynomial regression model: efficiency with weight restrictions ~ environmental variables 

```{r}

require(robustbase)
require(car)
require(stats)


# regression model with degree 2 to 6

# define degree 
degree <- 2

# go through loop with increasing degree
for (degree in 2:6){
  
  # polynomial regression model with feature selection 
poly <- lm(formula= mydata_balanced$Efficiency.with.Weight.restrictions ~ poly(mydata_balanced$Age.of.assets, degree) + poly(mydata_balanced$Average.declivity, degree) + poly(mydata_balanced$Intersection.with.indigenous.areas, degree) + poly(mydata_balanced$Height.of.vegetation, degree) + poly(mydata_balanced$Percentage.of.high.vegetation, degree) + poly(mydata_balanced$Average.precipitation, degree) + poly(mydata_balanced$Network.density, degree) + poly(mydata_balanced$Fire.incidence, degree) + poly(mydata_balanced$Number.of.intersections.with.roads, degree) + poly(mydata_balanced$Coincidence.of.network.lines,degree) + poly(mydata_balanced$Density.of.lightening,degree) + poly(mydata_balanced$Number.of.access.to.roads, degree), data = mydata_balanced)

# feature selection
poly <- step(poly, direction = "both", k=2, trace = 0)

# export console Output 
file <- file(paste("C:\\Users\\Kira\\OneDrive\\Controlling Seminar\\Seminar Controlling\\Abbildungen\\polynomial_regression_efficiency_with_wr_environmental_variables\\Summary_poly", degree, ".txt", sep=""))
sink(file, type = "output")

# summary and residual plot 
print(summary(poly))

# stop export 
sink()

# define filename for plot
filename <- paste("C:\\Users\\Kira\\OneDrive\\Controlling Seminar\\Seminar Controlling\\Abbildungen\\polynomial_regression_efficiency_with_wr_environmental_variables\\residualplot_", degree,".png", sep="")

# start export png
png(filename=filename, width=960, height=480)

# residual plot 
residualPlot(poly)

# stop export 
dev.off()

}

```
# chunk 4: polynomial regression degree 2 with feature selection
```{r}
require(robustbase)
require(car)
require(stats)

# define degree 
degree <- 2
  
  # polynomial regression model with feature selection 
poly_2 <- lm(formula= mydata_balanced$Efficiency.with.Weight.restrictions ~ poly(mydata_balanced$Age.of.assets, degree) + poly(mydata_balanced$Average.declivity, degree) + poly(mydata_balanced$Intersection.with.indigenous.areas, degree) +  poly(mydata_balanced$Percentage.of.high.vegetation, degree) + poly(mydata_balanced$Average.precipitation, degree) + poly(mydata_balanced$Network.density, degree) + poly(mydata_balanced$Fire.incidence, degree) + poly(mydata_balanced$Number.of.intersections.with.roads, degree) + poly(mydata_balanced$Coincidence.of.network.lines,degree) + poly(mydata_balanced$Density.of.lightening,degree) + poly(mydata_balanced$Number.of.access.to.roads, degree), data = mydata_balanced)

# feature selection
poly_2 <- step(poly_2, direction = "both", k=2, trace = 0)

# summary and residual plot 
print(summary(poly_2))

# residual plot 
residualPlot(poly_2)

# model prediction
yhat <- predict(poly_2)

# compute MSE and RMSE 
require(ModelMetrics)

MSE <- mse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("MSE")
print(MSE)

RMSE <- rmse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("RMSE")
print(RMSE)

# Breusch-Pagan-Test 
require(lmtest)

BP <- bptest(poly_2)

print(BP)

# Durbin-Watson Test

dwtest(formula = mydata_balanced$Efficiency.with.Weight.restrictions ~ 
    poly(mydata_balanced$Age.of.assets, degree) + poly(mydata_balanced$Average.declivity, 
        degree) + poly(mydata_balanced$Intersection.with.indigenous.areas, 
        degree) + poly(mydata_balanced$Percentage.of.high.vegetation, 
        degree) + poly(mydata_balanced$Average.precipitation, 
        degree) + poly(mydata_balanced$Network.density, degree) + 
        poly(mydata_balanced$Fire.incidence, degree) + poly(mydata_balanced$Number.of.intersections.with.roads, 
        degree), data = mydata_balanced)
```
# chunk 5:  cross validation poly_2
```{r}


require(caret)
set.seed(123)

# define degree
degree <- 2

# Define training control
train.control <- trainControl(method = "cv", number = 10)

# Train the model
 model <- train(x = data.frame(poly(mydata_balanced[, 13], degree), poly(mydata_balanced[, 15], degree), poly(mydata_balanced[, 16], degree), poly(mydata_balanced[, 18], degree), poly(mydata_balanced[, 19], degree), poly(mydata_balanced[, 20], degree), poly(mydata_balanced[, 21], degree), poly(mydata_balanced[, 22], degree)), y= mydata_balanced[,26], method = "lm", trControl = train.control)

 
# Summarize the results
 summary(model)
```
# chunk 6:  polynomial regression degree 3 with feature selection
```{r}
require(robustbase)
require(car)
require(stats)

# define degree 
degree <- 3
  
  # polynomial regression model with feature selection 
poly_3 <- lm(formula= mydata_balanced$Efficiency.with.Weight.restrictions ~ poly(mydata_balanced$Age.of.assets, degree) + poly(mydata_balanced$Average.declivity, degree) + poly(mydata_balanced$Intersection.with.indigenous.areas, degree) + poly(mydata_balanced$Height.of.vegetation, degree) + poly(mydata_balanced$Percentage.of.high.vegetation, degree) + poly(mydata_balanced$Average.precipitation, degree) + poly(mydata_balanced$Network.density, degree) + poly(mydata_balanced$Fire.incidence, degree) + poly(mydata_balanced$Number.of.intersections.with.roads, degree) + poly(mydata_balanced$Coincidence.of.network.lines,degree) + poly(mydata_balanced$Density.of.lightening,degree) + poly(mydata_balanced$Number.of.access.to.roads, degree), data = mydata_balanced)

# feature selection
poly_3 <- step(poly_3, direction = "both", k=2, trace = 0)

# summary and residual plot 
print(summary(poly_3))

# residual plot 
residualPlot(poly_3)

# model prediction
yhat <- predict(poly_3)

# compute MSE and RMSE
require(ModelMetrics)

MSE <- mse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("MSE")
print(MSE)

RMSE <- rmse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("RMSE")
print(RMSE)

```
# chunk 7: cross validation poly_3
```{r}


require(caret)
set.seed(123)

# define degree
degree <- 3

# Define training control
train.control <- trainControl(method = "cv", number = 10)

# Train the model
 model <- train(x = data.frame(poly(mydata_balanced[, 13], degree), poly(mydata_balanced[, 15], degree), poly(mydata_balanced[,16],degree), poly(mydata_balanced[, 17],degree), poly(mydata_balanced[, 19],degree)), y= mydata_balanced[,27], method = "lm", trControl = train.control)

 
# Summarize the results
 summary(model)
```
# chunk 8:  polynomial regression degree 4 with feature selection
```{r}
require(robustbase)
require(car)
require(stats)

# define degree 
degree <- 4
  
  # polynomial regression model with feature selection 
poly_4 <- lm(formula= mydata_balanced$Efficiency.with.Weight.restrictions ~ poly(mydata_balanced$Age.of.assets, degree) + poly(mydata_balanced$Average.declivity, degree) + poly(mydata_balanced$Intersection.with.indigenous.areas, degree) + poly(mydata_balanced$Height.of.vegetation, degree) + poly(mydata_balanced$Percentage.of.high.vegetation, degree) + poly(mydata_balanced$Average.precipitation, degree) + poly(mydata_balanced$Network.density, degree) + poly(mydata_balanced$Fire.incidence, degree) + poly(mydata_balanced$Number.of.intersections.with.roads, degree) + poly(mydata_balanced$Coincidence.of.network.lines,degree) + poly(mydata_balanced$Density.of.lightening,degree) + poly(mydata_balanced$Number.of.access.to.roads, degree), data = mydata_balanced)

# feature selection
poly_4 <- step(poly_4, direction = "both", k=2, trace = 0)

# summary and residual plot 
print(summary(poly_4))

# residual plot 
residualPlot(poly_4)

# model prediction
yhat <- predict(poly_4)

# compute MSE 
require(ModelMetrics)

MSE <- mse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("MSE")
print(MSE)

RMSE <- rmse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("RMSE")
print(RMSE)

```
# chunk 9: cross validation poly_4
```{r}


require(caret)
set.seed(123)

# define degree
degree <- 4

# Define training control
train.control <- trainControl(method = "cv", number = 10)

# Train the model
 model <- train(x = data.frame(poly(mydata_balanced[, 13], degree), poly(mydata_balanced[, 15], degree), poly(mydata_balanced[,16],degree), poly(mydata_balanced[, 18],degree)), y= mydata_balanced[,27], method = "lm", trControl = train.control)

 
# Summarize the results
 summary(model)
```
# chunk 10: polynomial regression degree 5 with feature selection
```{r}
require(robustbase)
require(car)
require(stats)

# define degree 
degree <- 5
  
  # polynomial regression model with feature selection 
poly_5 <- lm(formula= mydata_balanced$Efficiency.with.Weight.restrictions ~ poly(mydata_balanced$Age.of.assets, degree) + poly(mydata_balanced$Average.declivity, degree) + poly(mydata_balanced$Intersection.with.indigenous.areas, degree) + poly(mydata_balanced$Height.of.vegetation, degree) + poly(mydata_balanced$Percentage.of.high.vegetation, degree) + poly(mydata_balanced$Average.precipitation, degree) + poly(mydata_balanced$Network.density, degree) + poly(mydata_balanced$Fire.incidence, degree) + poly(mydata_balanced$Number.of.intersections.with.roads, degree) + poly(mydata_balanced$Coincidence.of.network.lines,degree) + poly(mydata_balanced$Density.of.lightening,degree) + poly(mydata_balanced$Number.of.access.to.roads, degree), data = mydata_balanced)

# feature selection
poly_5 <- step(poly_5, direction = "both", k=2, trace = 0)

# summary and residual plot 
print(summary(poly_5))

# residual plot 
residualPlot(poly_5)

# model prediction
yhat <- predict(poly_5)

# compute MSE and RMSE
require(ModelMetrics)

MSE <- mse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("MSE")
print(MSE)

RMSE <- rmse(mydata_balanced$Efficiency.with.Weight.restrictions, yhat)
print("RMSE")
print(RMSE)

```
# chunk 11:  cross validation poly_5
```{r}


require(caret)
set.seed(123)

# define degree
degree <- 5

# Define training control
train.control <- trainControl(method = "cv", number = 10)

# Train the model
 model <- train(x = data.frame(poly(mydata_balanced[, 13], degree), poly(mydata_balanced[, 15], degree), poly(mydata_balanced[,16],degree), poly(mydata_balanced[, 17],degree)), y= mydata_balanced[,27], method = "lm", trControl = train.control)

 
# Summarize the results
 summary(model)
 
 # Warning in predict.lm(modelFit, newdata) :
 # Vorhersage durch Fit ohne vollen Rang mag täuschen
```
